---
title: "Exploration of Stanford Open Policing Dataset"
author: "Linni Qin n9632981"
date: "4 March 2018"
output: html_document
---

## Group the offenders' age in groups as per the census formate in 3 states(WASHINGTON, COLORADO, NEVADA): 

# Calling or installing  the useful packages

```{r}
library(ggplot2)
library(plyr)
library(dplyr) # summarize
library(lubridate) # convert the date into date format "ymd" and for dealing with other operation on date
```

```{r,fig.width=10, fig.height=5}
# import the cleaned file and seperated violation text
co <- read.csv("./coseperated.csv", header = TRUE) 
nd <- read.csv("./ndseperated.csv", header = TRUE) 
nv <- read.csv("./nvseperated.csv", header = TRUE)  
wa <- read.csv("./waseperated.csv", header = TRUE)  
wy <- read.csv("./wyseperated.csv", header = TRUE)  

```

```{r,fig.width=10, fig.height=5}
co$stopDate <- ymd(co$stopDate) #convert the class of data "factor" into right format "date" for subset operation
nv$stopDate <- ymd(nv$stopDate) 
nd$stopDate <- ymd(nd$stopDate) 
wa$stopDate <- ymd(wa$stopDate) 
wy$stopDate <- ymd(wy$stopDate) 
```

# Explor the range of month each year in each state ******* (ONE TIME ONLY, SKIP THIS STEP TO PROCEED THE YEAR-LEVEL EXPLORATION)**********

```{r,fig.width=10, fig.height=5}
co$stopDate<-format(co$stopDate,"%Y-%m") #(2010-2016)
#unique(co$stopDate) 
```

```{r,fig.width=10, fig.height=5}
nd$stopDate<-format(nd$stopDate,"%Y-%m") #2010 - 2015
#unique(nd$stopDate)
```

```{r,fig.width=10, fig.height=5}
nv$stopDate<-format(nv$stopDate,"%Y-%m") # 2012-2016
#unique(nv$stopDate)
```

```{r,fig.width=10, fig.height=5}
wa$stopDate<-format(wa$stopDate,"%Y-%m") # 2009-2016
#unique(wa$stopDate)
```

```{r,fig.width=10, fig.height=5}
wy$stopDate<-format(wy$stopDate,"%Y-%m") # 2011-2012
#unique(wy$stopDate)
```
# *****************************************************************************************************************************************

# Explor year-level total amount of violation, by state

# convert all the stopDate into year only
```{r,fig.width=10, fig.height=5}
co$stopDate <- year(co$stopDate)
nv$stopDate <- year(nv$stopDate)
nd$stopDate <- year(nd$stopDate)
wa$stopDate <- year(wa$stopDate)
wy$stopDate <- year(wy$stopDate)

```

# Colorado State
```{r,fig.width=15, fig.height=5}
# get the count of violation based on stopDate
coViolationEachYear <- summarise(group_by(co,stopDate,violation),count =n())
#write.csv(coViolationEachYear, file="coViolationEachYear.csv", row.names = FALSE)

coViolationEachYear$stopDate <- as.character(coViolationEachYear$stopDate) 

ggplot(coViolationEachYear, aes(x=violation, y=count/1000, fill=stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 200), breaks = seq(0, 200, 25))+ ylab("count in thousand") +ggtitle("Colorado Traffic Violation Amount, by Year (2010 - 2016)")+  scale_fill_hue(name="Year")  +theme_bw()

```

```{r,fig.width=15, fig.height=5}
# test on percentage presentation in y axis

ggplot(coViolationEachYear, aes(x=violation, group=stopDate)) + geom_bar(aes(y=..prop.., fill = factor(..x..)), color="black", stat="count") + facet_grid(~stopDate)+ scale_y_continuous(labels=scales::percent)

```





```{r,fig.width=15, fig.height=5}
# Colorado State
# get the count of violation based on stopDate
coSpeed <- subset(coViolationEachYear, coViolationEachYear$violation == "Speeding")
coDUI <- subset(coViolationEachYear, coViolationEachYear$violation == "DUI")
coPhone <- subset(coViolationEachYear, coViolationEachYear$violation =="Cell phone")
coSafeMove <- subset(coViolationEachYear, coViolationEachYear$violation =="Safe movement")
coMoveVilation <- subset(coViolationEachYear, coViolationEachYear$violation =="Moving violation")

co5ViolationEachYear <- rbind(coSpeed, coSafeMove, coDUI, coPhone, coMoveVilation)
#write.csv(co5ViolationEachYear, file="co5ViolationEachYear.csv", row.names = FALSE)

co3ViolationEachYear <- rbind(coDUI, coPhone, coMoveVilation)

ggplot(co3ViolationEachYear, aes(x=violation, y=count/1000, fill=stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, 1))+ ylab("count in thousand") +ggtitle("Colorado Traffic Violation Amount, by Year, Only DUI, Phone, Moving Violation (2010 - 2016)")+  scale_fill_hue(name="Year")  +theme_bw()

```

# North Dakota State

```{r,fig.width=15, fig.height=5}
# get the count of violation based on stopDate
ndViolationEachYear <- summarise(group_by(nd,stopDate,violation),count =n())
#write.csv(ndViolationEachYear, file="ndViolationEachYear.csv", row.names = FALSE)

ndViolationEachYear$stopDate <- as.character(ndViolationEachYear$stopDate) 

ggplot(ndViolationEachYear, aes(x=violation, y=count/1000, fill=stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 45), breaks = seq(0, 45, 5)) + ylab("count in thousand") +ggtitle("North Dakota Violation Amount, by Year (2010 - 2015)")+  scale_fill_hue(name="Year")  +theme_bw()

```

```{r,fig.width=15, fig.height=5}

ndSpeed <- subset(ndViolationEachYear, ndViolationEachYear$violation == "Speeding")
ndDUI <- subset(ndViolationEachYear, ndViolationEachYear$violation == "DUI")
ndPhone <- subset(ndViolationEachYear, ndViolationEachYear$violation =="Cell phone")
ndSafeMove <- subset(ndViolationEachYear, ndViolationEachYear$violation =="Safe movement")
ndMoveVilation <- subset(ndViolationEachYear, ndViolationEachYear$violation =="Moving violation")

nd5ViolationEachYear <- rbind(ndSpeed, ndSafeMove, ndDUI, ndPhone, ndMoveVilation)
#write.csv(nd5ViolationEachYear, file="nd5ViolationEachYear.csv", row.names = FALSE)

nd3ViolationEachYear <- rbind(ndDUI, ndPhone,ndMoveVilation)

ggplot(nd3ViolationEachYear, aes(x=violation, y=count/1000, fill=stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 2), breaks = seq(0, 2, 0.2))+ ylab("count in thousand") +ggtitle("North Dakota Traffic Violation Amount, by Year, Only DUI, Phone (2010 - 2015)")+  scale_fill_hue(name="Year")  +theme_bw()

```

# Nevada State

```{r,fig.width=15, fig.height=5}

# get the count of violation based on stopDate
nvViolationEachYear <- summarise(group_by(nv,stopDate,violation),count =n())
#write.csv(nvViolationEachYear, file="nvViolationEachYear.csv", row.names = FALSE)

nvViolationEachYear$stopDate <- as.character(nvViolationEachYear$stopDate) 

ggplot(nvViolationEachYear, aes(x=violation, y=count/1000, fill=stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 25)) + ylab("count in thousand") +ggtitle("Nevada Traffic Violation Amount, by Year (2012 - 2016)")+  scale_fill_hue(name="Year")  +theme_bw()

```

```{r,fig.width=15, fig.height=5}
# Nevada State
nvSpeed <- subset(nvViolationEachYear, nvViolationEachYear$violation == "Speeding")
nvDUI <- subset(nvViolationEachYear, nvViolationEachYear$violation == "DUI")
nvPhone <- subset(nvViolationEachYear, nvViolationEachYear$violation =="Cell phone")
nvSafeMove <- subset(nvViolationEachYear, nvViolationEachYear$violation =="Safe movement")
nvMoveVilation <- subset(nvViolationEachYear, nvViolationEachYear$violation =="Moving violation")

nv5ViolationEachYear <- rbind(nvSpeed, nvSafeMove, nvDUI, nvPhone, nvMoveVilation)
#write.csv(nv5ViolationEachYear, file="nv5ViolationEachYear.csv", row.names = FALSE)

nv3ViolationEachYear <- rbind(nvDUI, nvPhone, nvMoveVilation)

ggplot(nv3ViolationEachYear, aes(x=violation, y=count/1000, fill=stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 18), breaks = seq(0, 18, 2))+ ylab("count in thousand") +ggtitle("Nevada Traffic Violation Amount, by Year, Only DUI, Phone, Moving violation (2012 - 2016)")+  scale_fill_hue(name="Year")  +theme_bw()

```


# Washington State

```{r,fig.width=15, fig.height=5}

# get the count of violation based on stopDate
waViolationEachYear <- summarise(group_by(wa,stopDate,violation),count =n())
#write.csv(waViolationEachYear, file="waViolationEachYear.csv", row.names = FALSE)

waViolationEachYear$stopDate <- as.character(waViolationEachYear$stopDate) # convert the numeric (2010 and 2016) to character for the color purpose in plot 

ggplot(waViolationEachYear, aes(x=violation, y=count/1000, fill=stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 500), breaks = seq(0, 500, 25))+ ylab("count in thousand") +ggtitle("Washington Traffic Violation Amount, by Year (2009 -2016)") +  scale_fill_hue(name="Year") +theme_bw()

```

```{r,fig.width=15, fig.height=5}
# Washington State
waSpeed <- subset(waViolationEachYear, waViolationEachYear$violation == "Speeding")
waDUI <- subset(waViolationEachYear, waViolationEachYear$violation == "DUI")
waPhone <- subset(waViolationEachYear, waViolationEachYear$violation =="Cell phone")
waSafeMove <- subset(waViolationEachYear, waViolationEachYear$violation =="Safe movement")
waMoveVilation <- subset(waViolationEachYear, waViolationEachYear$violation =="Moving violation")

wa5ViolationEachYear <- rbind(waSpeed, waSafeMove,waDUI, waPhone, waMoveVilation)
#write.csv(wa5ViolationEachYear, file="wa5ViolationEachYear.csv", row.names = FALSE)

wa3ViolationEachYear <- rbind(waDUI, waPhone, waMoveVilation)

ggplot(wa3ViolationEachYear, aes(x=violation, y=count/1000, fill=stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 25), breaks = seq(0, 25, 2))+ ylab("count in thousand") +ggtitle("Washington Traffic Violation Amount, by Year, Only DUI, Phone, Moving violation (2009 - 2016)")+  scale_fill_hue(name="Year")  +theme_bw()

```

# Wyoming State

```{r,fig.width=15, fig.height=5}

# get the count of violation based on stopDate
wyViolationEachYear <- summarise(group_by(wy,stopDate,violation),count =n())
#write.csv(wyViolationEachYear, file="wyViolationEachYear.csv", row.names = FALSE)

wyViolationEachYear$stopDate <- as.character(wyViolationEachYear$stopDate) 

ggplot(wyViolationEachYear, aes(x=violation, y=count/1000, fill=stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 70), breaks = seq(0, 70, 5))+ ylab("count in thousand") +ggtitle("Wyoming Traffic Violation Amount, by Year (2011-2012)") +  scale_fill_hue(name="Year") +theme_bw()

```


```{r,fig.width=15, fig.height=5}

wySpeed <- subset(wyViolationEachYear, wyViolationEachYear$violation == "Speeding")
wyDUI <- subset(wyViolationEachYear, wyViolationEachYear$violation == "DUI")
wyPhone <- subset(wyViolationEachYear, wyViolationEachYear$violation =="Cell phone")
wySafeMove <- subset(wyViolationEachYear, wyViolationEachYear$violation =="Safe movement")
wyMoveVilation <- subset(wyViolationEachYear, wyViolationEachYear$violation =="Moving violation")

wy5ViolationEachYear <- rbind(wySpeed, wySafeMove, wyDUI, wyPhone, wyMoveVilation)
#write.csv(wy5ViolationEachYear, file="wy5ViolationEachYear.csv", row.names = FALSE)

wy3ViolationEachYear <- rbind(wyDUI, wyPhone, wyMoveVilation)

ggplot(wy3ViolationEachYear, aes(x=violation, y=count/1000, fill=stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 2), breaks = seq(0, 2, 0.1))+ ylab("count in thousand") +ggtitle("Wyoming Traffic Violation Amount, by Year, Only DUI, Phone, Moving violation (2011 - 2012)")+  scale_fill_hue(name="Year")  +theme_bw()

```



## Washington State (Driver Census vs. Offenders Distribution, by Age group)

# plot the lines of wacensus

```{r,fig.width=10, fig.height=5}
wacensus <- read.csv("./wacensus.csv", header = TRUE)  
wacensus$year <- as.character(wacensus$year) # convert the numeric (2010 and 2016) to character for the color purpose in plot 
class(wacensus$year)

ggplot(wacensus, aes(x=driverAge, y=driverPopulation/1000, group = year)) + geom_line(aes(linetype=year))+ geom_point(aes(shape=year)) + ylab("driverPopulation in thousands")

library(scales)  # for percentage scales
ggplot(wacensus, aes(x=driverAge, y=driverPopulation/1000, fill=year)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))+ ylab("Thousands in Driver Population") +ggtitle("Washington Licenced Driver Population, by Age Group, by Year (2009 - 2015)")  +theme_bw()

```


# convert the age of driver into groups 

```{r,fig.width=10, fig.height=5}

wa$driverAge <- replace(wa$driverAge, wa$driverAge <= 19, "15-19")
wa$driverAge <- replace(wa$driverAge, wa$driverAge >=20 & wa$driverAge < 25, "20-24")
wa$driverAge <- replace(wa$driverAge, wa$driverAge >=25 & wa$driverAge< 30, "25-29")
wa$driverAge <- replace(wa$driverAge, wa$driverAge >=30 & wa$driverAge < 35, "30-34")
wa$driverAge <- replace(wa$driverAge, wa$driverAge >=35 & wa$driverAge < 40, "35-39")
wa$driverAge <- replace(wa$driverAge, wa$driverAge >=40 & wa$driverAge < 45, "40-44")
wa$driverAge <- replace(wa$driverAge, wa$driverAge >=45 & wa$driverAge < 50, "45-49")
wa$driverAge <- replace(wa$driverAge, wa$driverAge >=50 & wa$driverAge < 55, "50-54")
wa$driverAge <- replace(wa$driverAge, wa$driverAge >=55 & wa$driverAge < 60, "55-59")
wa$driverAge <- replace(wa$driverAge, wa$driverAge >=60 & wa$driverAge < 65, "60-64")
wa$driverAge <- replace(wa$driverAge, wa$driverAge >= 65, "65+")

```

# *************************************************SKIP IT IF THE YEAR HAS BEEN CONVERTED IN THE ABOVE CHUNCKS**********************
```{r,fig.width=18, fig.height=10}
# Extract the same year range with the census data (2009, 2015)
# The oplicing data covers 2010 (12 months)-2016 (1-3month)

wa$stopDate <- ymd(wa$stopDate) #convert the class of data "factor" into right format "date" for subset operation

wa09<-subset(wa, wa$stopDate >= "2009-01-01" & wa$stopDate < "2010-01-01")
wa09$stopDate <- c("2009")

wa15<-subset(wa, wa$stopDate >= "2015-01-01"& wa$stopDate < "2016-01-01")
wa15$stopDate <- c("2015")

wa0915 <- rbind(wa09, wa15)

```
# ********************************************************************************************************************************************
```{r,fig.width=10, fig.height=5}

# get the frequency count of driverAge based on stopDate
waAge <- summarise(group_by(wa,stopDate,driverAge),count =n())
write.csv(waAge, file="waOffender.csv", row.names = FALSE)

waAge$stopDate <- as.character(waAge$stopDate) 

ggplot(waAge, aes(x=driverAge, y=count/1000, fill = stopDate)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) + scale_y_continuous(limits = c(0, 240), breaks = seq(0, 240, 20))+ ylab("count in thousand") + ggtitle("Washington driver in groups and their violation frequency (2009 - 2016)") +  scale_fill_hue(name="Year")  +theme_bw()

```

## Colorado State (Driver Census vs. Offenders Distribution, by Age group)

# plot the lines of cocensus

```{r,fig.width=10, fig.height=5}
cocensus <- read.csv("./cocensus.csv", header = TRUE)  
cocensus$year <- as.character(cocensus$year) # convert the numeric (2010 and 2016) to character for the color purpose in plot 

ggplot(cocensus, aes(x=driverAge, y=driverPopulation/1000, group = year)) + geom_line(aes(linetype=year))+ geom_point(aes(shape=year))+ ylab("driverPopulation in thousands")

```

# convert the age of driver into groups 

```{r,fig.width=10, fig.height=5}

co$driverAge <- replace(co$driverAge, co$driverAge <= 19, "15-19")
co$driverAge <- replace(co$driverAge, co$driverAge >=20 & co$driverAge < 25, "20-24")
co$driverAge <- replace(co$driverAge, co$driverAge >=25 & co$driverAge< 30, "25-29")
co$driverAge <- replace(co$driverAge, co$driverAge >=30 & co$driverAge < 35, "30-34")
co$driverAge <- replace(co$driverAge, co$driverAge >=35 & co$driverAge < 40, "35-39")
co$driverAge <- replace(co$driverAge, co$driverAge >=40 & co$driverAge < 45, "40-44")
co$driverAge <- replace(co$driverAge, co$driverAge >=45 & co$driverAge < 50, "45-49")
co$driverAge <- replace(co$driverAge, co$driverAge >=50 & co$driverAge < 55, "50-54")
co$driverAge <- replace(co$driverAge, co$driverAge >=55 & co$driverAge < 60, "55-59")
co$driverAge <- replace(co$driverAge, co$driverAge >=60 & co$driverAge < 65, "60-64")
co$driverAge <- replace(co$driverAge, co$driverAge >= 65, "65+")

```

# Colorado State Traffic Offender distribution by age groups

```{r,fig.width=18, fig.height=10}
# Extract the same year range with the census data (2010, 2015)
# The oplicing data covers 2010 (12 months)-2016 (1-3month, and a few lines of other months)

co$stopDate <- ymd(co$stopDate) #convert the class of data "factor" into right format "date" for subset operation

co2010<-subset(co, co$stopDate >= "2010-01-01" & co$stopDate < "2011-01-01")
co2010$stopDate <- c("2010")

co2015<-subset(co, co$stopDate >= "2015-01-01" & co$stopDate < "2016-01-01")
co2015$stopDate <- c("2015")

co20102015 <- rbind(co2010, co2015)

```

```{r,fig.width=10, fig.height=5}

# get the frequency count of driverAge based on stopDate
coAge <- summarise(group_by(co20102015,stopDate,driverAge),count =n())
#write.csv(coAge, file="coOffender.csv", row.names = FALSE)

ggplot(coAge, aes(x=driverAge, y=count/1000, group = stopDate)) + geom_line(aes(linetype=stopDate))+ geom_point(aes(shape=stopDate))+ ylab("Offenders in thousands")

```

## Nevada State (Driver Census vs. Offenders Distribution, by Age group)

# plot the lines of necensus

```{r,fig.width=10, fig.height=5}
nvcensus <- read.csv("./nvcensus.csv", header = TRUE)  
nvcensus$year <- as.character(nvcensus$year) # convert the numeric (2010 and 2016) to character for the color purpose in plot 

ggplot(nvcensus, aes(x=driverAge, y=driverPopulation/1000, group = year)) + geom_line(aes(linetype=year))+ geom_point(aes(shape=year))+ ylab("driverPopulation in thousands")

```


# convert the age of driver into groups 

```{r,fig.width=10, fig.height=5}

nv$driverAge <- replace(nv$driverAge, nv$driverAge <= 19, "15-19")
nv$driverAge <- replace(nv$driverAge, nv$driverAge >=20 & nv$driverAge < 25, "20-24")
nv$driverAge <- replace(nv$driverAge, nv$driverAge >=25 & nv$driverAge< 30, "25-29")
nv$driverAge <- replace(nv$driverAge, nv$driverAge >=30 & nv$driverAge < 35, "30-34")
nv$driverAge <- replace(nv$driverAge, nv$driverAge >=35 & nv$driverAge < 40, "35-39")
nv$driverAge <- replace(nv$driverAge, nv$driverAge >=40 & nv$driverAge < 45, "40-44")
nv$driverAge <- replace(nv$driverAge, nv$driverAge >=45 & nv$driverAge < 50, "45-49")
nv$driverAge <- replace(nv$driverAge, nv$driverAge >=50 & nv$driverAge < 55, "50-54")
nv$driverAge <- replace(nv$driverAge, nv$driverAge >=55 & nv$driverAge < 60, "55-59")
nv$driverAge <- replace(nv$driverAge, nv$driverAge >=60 & nv$driverAge < 65, "60-64")
nv$driverAge <- replace(nv$driverAge, nv$driverAge >= 65, "65+")

```

# Nevada State Traffic Offender distribution by age groups

```{r,fig.width=18, fig.height=10}
# Extract the same year range with the census data (2013, 2015)
# The oplicing data covers 2012 (11 months from mid of Feb)-2016 (a couple of months)

nv$stopDate <- ymd(nv$stopDate) #convert the class of data "factor" into right format "date" for subset operation

nv13<-subset(nv, nv$stopDate >= "2013-01-01" & nv$stopDate < "2014-01-01")
nv13$stopDate <- c("2013")

nv15<-subset(nv, nv$stopDate >= "2015-01-01" & nv$stopDate < "2016-01-01")
nv15$stopDate <- c("2015")

nv1315 <- rbind(nv13, nv15)

```

```{r,fig.width=10, fig.height=5}

# get the frequency count of driverAge based on stopDate
nvAge <- summarise(group_by(nv1315,stopDate,driverAge),count =n())
#write.csv(nvAge, file="nvOffender.csv", row.names = FALSE)

ggplot(nvAge, aes(x=driverAge, y=count/1000, group = stopDate)) + geom_line(aes(linetype=stopDate))+ geom_point(aes(shape=stopDate))+ ylab("Offenders in thousands")

```


## Explore the violation distribution

# Washington state 2009

```{r,fig.width=20, fig.height=15}

# get the frequency count of violation based on driverAge
waV09 <- summarise(group_by(wa09,driverAge,violation),count =n())

ggplot(waV09, aes(x=violation, y=count, fill=driverAge)) + geom_bar(stat = "identity") # no obvious patterns to tell the insight of dui, cellphone, speeding within the varied age groups

```

```{r,fig.width=20, fig.height=10}

# targetting at speeding, dui, cell phone and age groups)

waSpeed09 <- subset(waV09, waV09$violation == "Speeding")
waDUI09 <- subset(waV09, waV09$violation == "DUI")
waPhone09 <- subset(waV09, waV09$violation == "Cell phone")

wa3V09 <- rbind(waSpeed09, waDUI09, waPhone09)
#write.csv(wa3V09, file="wa3violation09.csv", row.names = FALSE)

ggplot(wa3V09, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Washington Traffice Violation and Age Groups (2009)") +theme_bw()

```




```{r,fig.width=20, fig.height=10}

# as the number of speeding violation is extraordinarily big, hard to investigate clearly the patterns of cell phone and dui, so I explore the latter two viloation individually for supplement.

wa2V09 <- rbind(waDUI09, waPhone09)
ggplot(wa2V09, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Washington Traffice Violation and Age Groups (2009) - DUI and Cell Phone only") +theme_bw()

```

# Washington state 2015

```{r,fig.width=20, fig.height=10}

# get the frequency count of violation based on driverAge
waV15 <- summarise(group_by(wa15,driverAge,violation),count =n())

# targetting at speeding, dui, cell phone and age groups)

waSpeed15 <- subset(waV15, waV15$violation == "Speeding")
waDUI15 <- subset(waV15, waV15$violation == "DUI")
waPhone15 <- subset(waV15, waV15$violation == "Cell phone")

wa3V15 <- rbind(waSpeed15, waDUI15, waPhone15)
#write.csv(wa3V15, file="wa3violation15.csv", row.names = FALSE)

ggplot(wa3V15, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Washington Traffice Violation and Age Groups (2015)") +theme_bw()


```

```{r,fig.width=20, fig.height=10}

# as the number of speeding violation is extraordinarily big, hard to investigate clearly the patterns of cell phone and dui, so I explore the latter two viloation individually for supplement.

wa2V15 <- rbind(waDUI15, waPhone15)
ggplot(wa2V15, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Washington Traffice Violation and Age Groups (2015) - DUI and Cell Phone only") +theme_bw()

```


# Colorado state 2010

```{r,fig.width=20, fig.height=10}

# get the frequency count of violation based on driverAge
coV10 <- summarise(group_by(co2010,driverAge,violation),count =n())

# targetting at speeding, dui, cell phone and age groups)

coSpeed10 <- subset(coV10, coV10$violation == "Speeding")
coDUI10 <- subset(coV10, coV10$violation == "DUI")
coPhone10 <- subset(coV10, coV10$violation == "Cell phone")

co3V10 <- rbind(coSpeed10, coDUI10, coPhone10)
#write.csv(co3V10, file="co3violation10.csv", row.names = FALSE)

ggplot(co3V10, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Colorado Traffice Violation and Age Groups (2010)") +theme_bw()


```
```{r,fig.width=20, fig.height=10}

co1V10 <- rbind(coDUI10)

ggplot(co1V10, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Colorado Traffice Violation and Age Groups (2015) - DUI only") +theme_bw()


```

# Colorado state 2015

```{r,fig.width=20, fig.height=10}

# get the frequency count of violation based on driverAge
coV15 <- summarise(group_by(co2015,driverAge,violation),count =n())

# targetting at speeding, dui, cell phone and age groups)

coSpeed15 <- subset(coV15, coV15$violation == "Speeding")
coDUI15 <- subset(coV15, coV15$violation == "DUI")
coPhone15 <- subset(coV15, coV15$violation == "Cell phone")

co3V15 <- rbind(coSpeed15, coDUI15, coPhone15)
#write.csv(co3V15, file="co3violation15.csv", row.names = FALSE)

ggplot(co3V15, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Colorado Traffice Violation and Age Groups (2015)") +theme_bw()


```


```{r,fig.width=20, fig.height=10}

co2V15 <- rbind(coDUI15, coPhone15)

ggplot(co2V15, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Colorado Traffice Violation and Age Groups (2015) - DUI and Cell Phone only") +theme_bw()


```



# Nevada state 2013

```{r,fig.width=20, fig.height=10}

# get the frequency count of violation based on driverAge
nvV13 <- summarise(group_by(nv13,driverAge,violation),count =n())

# targetting at speeding, dui, cell phone and age groups)

nvSpeed13 <- subset(nvV13, nvV13$violation == "Speeding")
nvDUI13 <- subset(nvV13, nvV13$violation == "DUI")
nvPhone13 <- subset(nvV13, nvV13$violation == "Cell phone")

nv3V13 <- rbind(nvSpeed13, nvDUI13, nvPhone13)
#write.csv(nv3V13, file="nv3violation13.csv", row.names = FALSE)

ggplot(nv3V13, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Nevada Traffice Violation and Age Groups (2013)") +theme_bw()


```

```{r,fig.width=20, fig.height=10}

nv2V13 <- rbind(nvDUI13, nvPhone13)

ggplot(nv2V13, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Nevada Traffice Violation and Age Groups (2013) - DUI and Cell Phone only") +theme_bw()


```



# Nevada state 2015

```{r,fig.width=20, fig.height=10}

# get the frequency count of violation based on driverAge
nvV15 <- summarise(group_by(nv15,driverAge,violation),count =n())

# targetting at speeding, dui, cell phone and age groups)

nvSpeed15 <- subset(nvV15, nvV15$violation == "Speeding")
nvDUI15 <- subset(nvV15, nvV15$violation == "DUI")
nvPhone15 <- subset(nvV15, nvV15$violation == "Cell phone")

nv3V15 <- rbind(nvSpeed15, nvDUI15, nvPhone15)
#write.csv(nv3V15, file="nv3violation15.csv", row.names = FALSE)

ggplot(nv3V15, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Nevada Traffice Violation and Age Groups (2015)") +theme_bw()


```


```{r,fig.width=20, fig.height=10}

nv2V15 <- rbind(nvDUI15, nvPhone15)

ggplot(nv2V15, aes(x=driverAge, y=count, fill=violation)) + geom_bar(color="black", stat="identity", position=position_dodge(), size=.4) +  ggtitle("Nevada Traffice Violation and Age Groups (2015) - DUI and Cell Phone only") +theme_bw()


```









































